---
title: "p8105_hw3_jl5548"
author: "J L"
date: "October 14, 2019"
output: github_document
---

```{r setup, include=FALSE}
library(ggridges)
library(patchwork)
library(tidyverse)
library(viridis)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## Problem 1

```{r}
## Load data and view
library(p8105.datasets)
data("instacart")
instacart
```

The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations. 

```{r, eval = FALSE}
## Understand variables
str(instacart)
nrow(instacart)
ncol(instacart)
```
The dataset instacart records online grocery orders from Instacart users and contains `r nrow(instacart)` observations of 131,209 unique users. Each row is a product from an order, and each user only place one order in this dataset. There are `r ncol(instacart)` variables in this dataset:
*order_id: order identifier
*product_id: product identifier
*add_to_cart_order: order in which each product was added to cart
*reordered: 1 if this prodcut has been ordered by this user in the past, 0 otherwise
*user_id: customer identifier
*eval_set: which evaluation set this order belongs in (Note that the data for use in this class is exclusively from the “train” eval_set)
*order_number: the order sequence number for this user (1=first, n=nth)
*order_dow: the day of the week on which the order was placed
*order_hour_of_day: the hour of the day on which the order was placed
*days_since_prior_order: days since the last order, capped at 30, NA if order_number=1
*product_name: name of the product
*aisle_id: aisle identifier
*department_id: department identifier
*aisle: the name of the aisle
*department: the name of the department
Among these variables, some of the key variables are reordered, order_dow, order_hour_of_day, days_since_prior_order, product_name, and aisle. The variable reordered informs us whether the product has been ordered in the past by the same user. Order_dow and order_hour_of_day informs us which day of the week and what hour the order was placed, respectively. Days_since_prior_order gives us a sense of the order cycle of a customer. Product_name and aisle provides us the information of which product and to which aisle the product belong, respectively. For instance, in the observation shown below, the product Bulgarian yogurt (product_name) from dairy/egg department (department) and yogurt aisle (aisle) was ordered on Thursday (order_dow) at 10am (order_hour_of_day), 9 days since the same person ordered it last time (days_since_prior_order). 
```{r}
instacart %>% filter(order_id == 1, product_id == 49302)
```


Then, do or answer the following (commenting on the results of each):

* How many aisles are there, and which aisles are the most items ordered from?
```{r}
## reorder the dataframe
df_aisle = instacart %>% 
  arrange(aisle_id, aisle) %>% 
  group_by(aisle_id, aisle) %>% 
  summarize(count = n()) 
most_ordered_aisle = df_aisle %>% 
  filter(count == max(pull(df_aisle, count))) %>% 
  select(aisle)
```
There are `r nrow(df_aisle)` aisles, and aisle #`r pull(most_ordered_aisle, aisle_id)`, `r pull(most_ordered_aisle, aisle)`, are the most items ordered from.

* Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r}
plot_aisle = df_aisle %>% 
  filter(count > 10000) %>% 
  ggplot(aes(x = count, y = aisle, size = count, color = count)) +
  geom_point()
plot_aisle
```
As shown in the plot above, y-axis contains all the aisle names and x-axis contains the corresponding number of items ordered in each aisle (those with more than 10000 items ordered). Number of items ordered in each aisle is represented by the size and color of the points, scale shown in the bottom of the graph. 

* Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
```{r}
most_popular = instacart %>% 
  filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") %>% 
  group_by(aisle, product_name) %>% 
  summarize(
    count = n(),
    time_ordered = sum(order_number)
    ) %>% 
  top_n(3, count) %>% 
  arrange(aisle, desc(count), product_name) 
knitr::kable(most_popular)
```

* Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).
```{r}
weekday = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")
mean_hour_ordered = instacart %>%
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>% 
  select(product_name, order_dow, order_hour_of_day) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(
    mean_hour = mean(order_hour_of_day)
  ) %>% 
  mutate(order_dow = weekday[order_dow + 1]) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) 
knitr::kable(mean_hour_ordered)
```



## Problem 2

Preparation
```{r}
## Load data and view
data("brfss_smart2010")
brfss_smart2010
```

```{r, eval = FALSE}
## Understand variables
str(brfss_smart2010)
nrow(brfss_smart2010)
ncol(brfss_smart2010)
```


First, do some data cleaning:

* format the data to use appropriate variable names
* focus on the “Overall Health” topic
* include only responses from “Excellent” to “Poor”
* organize responses as a factor taking levels ordered from “Poor” to “Excellent”

```{r}
df_brfss = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  separate(locationdesc, into = c("state", "county"), sep = "-") %>% 
  select(-locationabbr) %>% 
  filter(topic == "Overall Health") %>% 
  filter(response == "Poor" | response == "Fair" | response == "Good" | response == "Very Good" | response == "Excellent") %>% 
  mutate(response = factor(response, levels = c("Poor", "Fair", "Good", "Very Good", "Excellent")))
```


Using this dataset, do or answer the following (commenting on the results of each):

* In 2002, which states were observed at 7 or more locations? What about in 2010?
```{r}
state_2002 = df_brfss %>% 
  filter(year == 2002) %>% 
  group_by(state) %>% 
  distinct(county) %>% 
  summarize(num_location = n()) %>% 
  filter(num_location > 6) %>% 
  pull(state)

state_2010 = df_brfss %>% 
  filter(year == 2010) %>% 
  group_by(state) %>% 
  distinct(county) %>% 
  summarize(num_location = n()) %>% 
  filter(num_location > 6) %>% 
  pull(state)

state_2002
state_2010
```
As shown above, `r state_2002` were observed at 7 or more locations in 2002, and `r state_2010` were observed at 7 or more locations in 2010.

* Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).
```{r}

```

* Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.
```{r}

```



## Problem 3